Rigid fluid coupling

### 2022-10-17 manuscript/paper.tex
Add coupling equations. The code is messy now. But this is archieved

\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{relsize}
\usepackage[utf8]{inputenc}
\usepackage{cleveref}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage[section]{placeins}
\usepackage{booktabs}

@@

\section{Introduction}
\label{sec:intro}
\begin{itemize}
\item Rigid bodies in fluid flows in nature. Where can we find them? What kind
  of problems do they give rise to? Ice-Sea? Debris?
\item Why is it hard to model rigid fluid coupling? Then we say numerically
  nonlinear, and this is how it is been into handled. Many different ways RFC
  is been modeled, this includes mesh based and messless. Explain in a good
  detail what meshless methods are out there which handles these problems.
  This includes SPH too. (Same as next point)
\item \todoin{Done} Describe some works where Rigid fluid coupling is handled
  in meshless methods. MPS-DEM, SPH-DEM and other ways, recently Asai. (Same
  as previous point)
\item Explain SPH history and its applications in different areas. Now SPH
  specifically applied in RFC area. Detail some papers where how differently
  fluid is modelled, this includes ISPH, WCSPH, TVF, delta+ SPH. Also mention
  the latest multiphase flow, where air is considered, then mention the
  adaptive formulation, where the author captures the cavity.
\item \todoin{Done} Similarly explain the many ways the rigid-rigid coupling is been
  handled. Detail how non-smooth boundaries are handled.
\item \todoin{Done} Explain what we are going to do in the current work. We adapt the
  Mohseni formulation to handle the non-smooth boundaries, we also incorporate
  the damping, and extend the formulation to 3d and simulate 3d problems.
\item Explain the examples. 3d Sliding case, 3d controlled sliding case, body
  falling in water, then 3d dam break over dynamics rigid obstacles.
% \item Mention the open source software which provide SPH formulation (there are 2.)
\end{itemize}

@@

\textbf{1st bullet point of introduction}. Irregular bodies are modeled using
many different variations now, one of the methods is by overlapped sphere
method \cite{das2007modeling}, SMR-DEM by \citet{zhan2021surface}, dilated
polyhedral DEM \citet{liu2020new}, Fourier series-based discrete element
method \citet{lai2020fourier}, non-smooth wall modeling by
\citet{amaro2019improvement}, GJK-DEM, DFR-DEM. In the current work we have
extended the contact force model proposed by \citet{mohseni2021particle} to
handle three dimensional problems, we also incorporate the damping model.

\textbf{5th bullet point of introduction}. Irregular bodies are modeled using
many different variations now, one of the methods is by overlapped sphere
method \cite{das2007modeling}, SMR-DEM by \citet{zhan2021surface}, dilated
polyhedral DEM \citet{liu2020new}, Fourier series-based discrete element
method \citet{lai2020fourier}, non-smooth wall modeling by
\citet{amaro2019improvement}, GJK-DEM, DFR-DEM. In the current work we have
extended the contact force model proposed by \citet{mohseni2021particle} to
handle three dimensional problems, we also incorporate the damping model.

\textbf{6th bullet point of introduction}. Numerical examples, collision of
three cubes is initially used to test the conservation property of the current
contact mode while dealing with non-smooth boundaries. Further, a free and
forced sliding of a 3d cube on a frictional inclined plane, and a rolling
cylinder whose analytical solutions exists are utilised to validate the
rigid-rigid contact model frictional model. A granular collapse of a stack of
cylinders is modeled and compared against the experimental results. In order
to validate the two-way rigid fluid coupled scheme a rigid circular body
falling in a steady tank is simulated. Complex interactions of fluid in a 3D
dam breaking with multiple rigid bodies, is studied and compared with the
experimental results of SPH-DCDEM.

@@

The mass and momentum conservation laws that govern the dynamics of fluid motion are given in Lagrangian form by:
\begin{equation}
  \label{eq:ce}
  \frac{d \rho}{d t} = - \rho \; \frac{\partial u_i}{\partial x_i},
\end{equation}
\begin{equation}
  \label{eq:me}
  \frac{d u_i}{d t} = \frac{1}{\rho} \; \frac{\partial \sigma_{ij}}{\partial x_j}
  + g_i,
\end{equation}
where $\rho$ is the density, $u_i$ is the $i$\textsuperscript{th} component of
the velocity field, $x_j$ is the $j$\textsuperscript{th} component of the
position vector, $g_i$ is the component of body force per unit mass. And $\frac{d}{d t}$ is
\begin{equation}
  \label{eq:material-derivative}
  \frac{d}{d t} = \frac{\partial }{\partial t} +
  {u}_j \frac{\partial }{\partial x_j}.
\end{equation}

Following CTVF formulation, and moving the particles with transport velocity,
we have the modified material derivative as
\begin{equation}
  \label{eq:modified-material-derivative}
  \frac{\tilde{d} }{d t} = \frac{\partial }{\partial t} +
  \tilde{u}_j \frac{\partial }{\partial x_j}.
\end{equation}

The adjust equations for the modified derivative are given by:

\begin{equation}
  \label{eq:ce-tvf}
  \frac{\tilde{d} \rho}{d t} =
  - \rho \frac{\partial \tilde{u}_j}{\partial x_j} +
  \frac{\partial (\rho (\tilde{u}_j - u_j))}{\partial x_j}.
\end{equation}
\begin{equation}
  \label{eq:mom-tvf}
  \frac{\tilde{d} u_i}{d t} =
  \frac{\partial}{\partial x_j} (u_i (\tilde{u}_j - u_j))
  - u_i \frac{\partial}{\partial x_j} (\tilde{u}_j - u_j)
  + g_i
  +\frac{1}{\rho} \frac{\partial \sigma_{ij}}{\partial x_j}.
\end{equation}

For a weakly-compressible or incompressible fluid, a viscous force is added:
\begin{equation}
  \label{eq:fluid-stress-decomposition}
  \sigma_{ij} = - p \delta_{ij} + 2 \eta \frac{\partial u_i}{\partial x_j}
\end{equation}
where $\eta$ is the kinematic viscosity of the fluid.

In both fluid and solid modelling the pressure is computed using an
isothermal equation of state, given as,
\begin{equation}
  \label{eq:pressure-equation}
  p = K \bigg(\frac{\rho}{\rho_{0}} - 1 \bigg),
\end{equation}
where $K = \rho_{0} c_0^2$ is the bulk modulus. Here, the constants $c_0$ and
$\rho_0$ are the reference speed of sound and density, respectively. For solids,
$c_0$ is computed as $\sqrt{\frac{E}{3 (1 - 2 \nu)\rho_{0}}}$, $\nu$ is the
Poisson ratio.

SPH discretization of the modified governing equations are given by:
\begin{equation}
  \label{eq:sph-discretization-continuity}
  \frac{\tilde{d}\rho_a}{dt} = \sum_{b} \; \frac{m_b}{\rho_{b}} \; (
  \rho_{a} \; \tilde{\ten{u}}_{ab} \; + \;
  (\rho \; (\tilde{\ten{u}} \; - \;
  \ten{u}))_{ab}) \; \cdot \nabla_{a} W_{ab},
\end{equation}
\begin{multline}
  \label{eq:sph-momentum-fluid}
  \frac{\tilde{d}\ten{u}_{a}}{dt} = - \sum_{b} m_b \bigg[
  \bigg(\frac{p_a}{\rho_a^2} + \frac{p_b}{\rho_b^2}\bigg) \ten{I} -
  \bigg(\frac{\ten{A}_a}{\rho_a^2} + \frac{\ten{A}_b}{\rho_b^2}
  \bigg) \bigg]
  \cdot \nabla_{a} W_{ab} \\
  + \ten{u}_{a} \sum_{b} \frac{m_b}{\rho_{b}} \; \tilde{\ten{u}}_{ab} \cdot
  \nabla_{a} W_{ab} + \sum_{b} m_b \frac{4 \eta \nabla W_{ab}\cdot
    \ten{r}_{ab}}{(\rho_a + \rho_b) (r_{ab}^2 + 0.01 h_{ab}^2)} \ten{u}_{ab} +
  \ten{g}_{a},
\end{multline}
The particles in the current scheme are moved with the transport velocity,
\begin{equation}
  \label{eq:transport_velocity_position_derivative}
  \frac{d\ten{r}_a}{dt} = \ten{\tilde{u}}_a.
\end{equation}
%
The transport velocity is updated using,
\begin{equation}
  \label{eq:transport_velocity}
  \ten{\tilde{u}}_a(t + \Delta t) =\ten{u}_a(t) + \Delta t \; \frac{\tilde{d} \ten{u}_a}{dt} +
  \bigg(\frac{d \ten{u}_{a}}{dt}\bigg)_{\text{c}} \Delta t,
\end{equation}
where $\ten{A}_a = \rho_a \ten{u}_a \otimes (\ten{\tilde{u}}_a - \ten{u}_a)$,
$\ten{I}$ is the identity matrix, $\eta$ is the kinematic viscosity of the
fluid and \citet{morris1997modeling} formulation is used to discretize the
viscosity term.
\begin{itemize}
\item Discuss about the homogenization force (We use GTVf).
\item Describe $\Pi_{ij}$ for numerical stability.
\item And other terms.
\end{itemize}
We add to the momentum equation an additional artificial viscosity term
$\Pi_{ab}$~\cite{monaghan-review:2005} to maintain the stability of the
numerical scheme, given as,
\begin{align}
  \label{eq:mom-av}
  \Pi_{ab} =
  \begin{cases}
\frac{-\alpha h_{ab} \bar{c}_{ab} \phi_{ab}}{\bar{\rho}_{ab}}
  & \ten{u}_{ab}\cdot \ten{r}_{ab} < 0, \\
  0 & \ten{u}_{ab}\cdot \ten{r}_{ab} \ge 0,
\end{cases}
\end{align}
where,
%
\begin{equation}
  \label{eq:av-phiij}
  \phi_{ab} = \frac{\ten{u}_{ab} \cdot \ten{r}_{ab}}{r^2_{ab} + 0.01 h^2_{ab}},
\end{equation}
%
where $\ten{r}_{ab} = \ten{r}_a - \ten{r}_b$, $\ten{u}_{ab} = \ten{u}_a -
\ten{u}_b$, $h_{ab} = (h_a + h_b)/2$, $\bar{\rho}_{ab} = (\rho_a + \rho_b)/2$,
$\bar{c}_{ab} = (c_a + c_b) / 2$, and $\alpha$ is the artificial
viscosity parameter.

%
where $\big(\frac{d \ten{u}_a}{dt}\big)_{\text{c}}$ is the homogenization
acceleration which ensures that the particle positions are homogeneous.

\subsection{Boundary conditions}

The ghost particle approach of \citet{Adami2012} is used to model the
boundaries. We use three layers of ghost particles to model the solid wall.
The properties of the solid wall are interpolated from the fluid particles.

When computing the divergence of the velocity field on fluid particles, we
enforce a no-penetration boundary condition and not a no-slip boundary
condition. The velocity of the fluid is projected onto the ghost particles
using,
\begin{equation}
  \label{eq:v-ghost}
  \ten{\hat{u}}_a = \frac{\sum_b\ten{u}_b W_{ab}}{\sum_b W_{ab}},
\end{equation}
\begin{equation}
  \label{eq:v-hat-ghost}
  \ten{\check{u}}_a = \frac{\sum_b\tilde{\ten{u}}_b W_{ab}}{\sum_b W_{ab}},
\end{equation}
where $\ten{u}_b$, $\ten{\tilde{u}}_b$ are the momentum and transport velocity
of the fluid respectively and $W_{ab}$ is the kernel value between the fluid
particle and the ghost particle.

The normal component of this projected velocity is then reflected and set as
the ghost particle velocity,
\begin{equation}
  \label{eq:free-slip-bc-u}
  \ten{u}_{\text{Ga}} = 2 \ten{\hat{n}}((\ten{u}_{\text{p}} - \ten{\hat{u}}_{\text{a}})\cdot \ten{\hat{n}}) + \ten{\hat{u}}_{\text{a}},
\end{equation}
where $\ten{u}_{\text{p}}$ is the local velocity of the boundary and
$\ten{\hat{n}}$ is the unit normal to the boundary particle $a$. Similarly the
transport velocity of the ghost particle is set as,
\begin{equation}
  \label{eq:free-slip-bc-u}
  \tilde{\ten{u}}_{\text{Gi}} = 2 \ten{\hat{n}}((\ten{u}_{\text{p}} - \ten{\check{u}}_{\text{i}})\cdot \ten{\hat{n}}) + \ten{\check{u}}_{\text{i}},
\end{equation}

When the viscous force is computed, the no slip boundary condition is used,
where the velocity on the boundary set as,
\begin{equation}
  \label{eq:no-slip-bc-u}
  \ten{u}_{\text{Ga}} = 2 \ten{u}_{\text{p}} - \ten{\hat{u}}_{\text{a}},
\end{equation}
a similar form is used for the transport velocity here too,
\begin{equation}
  \label{eq:no-slip-bc-uhat}
  \tilde{\ten{u}}_{\text{Ga}} = 2 \ten{u}_{\text{p}} - \ten{\check{u}}_{\text{a}}.
\end{equation}

The pressure of the boundary particle is extrapolated from its surrounding
fluid particles by the following equation,
\begin{equation}
  \label{eq:pressure-bc}
  p_w = \frac{\Sigma_f p_f W_{wf} + (\ten{g} - \ten{a}_{\ten{w}}) \cdot \Sigma_f
    \rho_f \ten{r}_{wf} W_{wf}}{\Sigma_f W_{wf}},
\end{equation}
where $\ten{a}_w$ is the acceleration of the wall. The subscript $f$ denotes
the fluid particles and $w$ denotes the wall particles.

\begin{itemize}
\item Choose the timestep
\end{itemize}

@@

% check \cite{natsui2018sph}

The state of a rigid body is defined by position of center of mass, velocity
of center of mass, a rotation matrix to define the three rotational degrees of
freedom and angular velocity.

The equations governing the dynamics of rigid body are, balance of linear and
angular momentum given by,

\begin{equation}
  \label{eq:balance_linear_mom}
  \frac{d \; (M \ten{v}_{cm})}{d t} = \sum_i \ten{F}_i
\end{equation}

\begin{equation}
  \label{eq:balance_angular_mom}
  \frac{d \ten{L}}{d t} = \teng{\tau}_{cm}
\end{equation}

where $\teng{\tau}_i$ is the torque about center of mass due to forces on the
particles,

\begin{equation}
  \label{eq:torque}
  \teng{\tau}_{cm} = \sum_i \ten{F}_i \times (\ten{r}_{cm} - \ten{r}_{i})
\end{equation}

The angular momentum and the center of mass velocity of the body is updated by following

\begin{equation}
  \label{eq:ang_mom_update}
  \ten{x}_{cm} = \ten{x}_{cm} + \frac{\ten{F}_{cm}}{M} \; \Delta t
\end{equation}

\begin{equation}
  \label{eq:ang_mom_update}
  \ten{L} = \ten{L} + \teng{\tau}_{cm} \; \Delta t
\end{equation}

\begin{equation}
  \label{eq:rotation_update}
  \ten{R} = \ten{R} + \tilde{\teng{\omega}} \, \ten{R} \; \Delta t
\end{equation}

\begin{equation}
  \label{eq:moi_update}
  \ten{I}^{-1} = \ten{R} \hat{\ten{I}}^{-1} \ten{R}^T
\end{equation}

\begin{equation}
  \label{eq:moi_update}
  \teng{\omega} = \ten{I}^{-1} \; \ten{L}
\end{equation}

Particles update

@@

\begin{multline}
  \label{eq:force_on_fluid_due_to_rb}
  \frac{\tilde{d}\ten{u}_{a}}{dt} =  \sum_{b} m^0_b \bigg[
  \bigg(\frac{p_a}{\rho_a^2} + \frac{p_b}{\rho_b^2}\bigg) \ten{I}\bigg) \bigg]
  \cdot \nabla_{a} W_{ab}
\end{multline}

\begin{multline}
  \label{eq:force_on_fluid_due_to_rb}
  f_i =  m_a \sum_{b} m_b \bigg[
  \bigg(\frac{p_a}{\rho_a^2} + \frac{p_b}{\rho_b^2}\bigg) \ten{I}\bigg) \bigg]
  \cdot \nabla_{a} W_{ab}
\end{multline}

@@

\FloatBarrier%
\section{Collision between rigid bodies}
\label{sec:contact-force}

\citet{chen2019coupled} has damping model. Also he mentioned the parameters
required in stack of cylinders example.

>>>>

\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{relsize}
\usepackage[utf8]{inputenc}
\usepackage{cleveref}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage[section]{placeins}
\usepackage{booktabs}
\usepackage{physics}
\usepackage{breqn}

@@

\section{Introduction}
\label{sec:intro}

Arbitrary shaped rigid bodies in fluid flows in nature. Debris flowing,
particle transport chemical industries, food processing, in natural hazard
where the bodies are capable of doing catastrophic damage while being
sedimented. In ice-sea modeling. Study of such physics is hard to model, as
the two-way coupling is nonlinear. Many different models were proposed to
handle the current physics.

In meshless methods, MPS-DEM \cite{guo2017numerical} and SPH-DEM
\cite{canelas2016sph} are two ways of modeling the physics. Where Moving
Particle Semi-implicit (MPS) method and smoothed particle hydrodynamics (SPH)
are utilized to study the fluid dynamics, while the Discrete Element Method
(DEM) is used to model the interaction among the rigid bodies and finally both
SPH and MPS are coupled with the rigid bodies to handle the two way coupling.
In addition to DEM alone to model the rigid-rigid interaction, there are also
works where SPH \cite{amicarelli2015smoothed} itself is used to model the
interaction. But it lacks to incorporate the friction between the bodies. In
the current work we have used SPH to model the fluid dynamics.

The SPH method is a meshless numerical method originally proposed by
\citet{monaghan-gingold-stars-mnras-77} and \citet{lucy1977numerical} to model
astrophysics problems. It is fairly success in modeling physics ranging
different fields, \citet{monaghan-review:2005} provides a detailed review of
SPH and its applications. structural dynamics, FSI, fracture to note a few.
SPH has been extensively applied to model the dynamics of fluid in two way
rigid fluid coupling. There are several SPH variants in which the fluid flow
can be modeled. \citet{canelas2016sph} uses WCSPH in rigid fluid coupling in
modeling two way rigid fluid coupling, ISPH is used in \cite{asai2021fluid},
delta sph in \cite{lyu20223d}. Multiphase flow, where air is considered, to
capture the cavity while the rigid body enters the water section is modeled in
\cite{lyu2021study}. A TVF scheme is currently proposed where the particle
shifting is introduced for better particle distribution leading to less errors
is not yet utilized in modeling the rigid fluid coupling.

All the works above are using DEM to model the rigid-rigid interaction doesn't
consider the shape of the body and fails to model bodies with arbitrary shape.
One of the reasons is the contact force model is between two particles, and
those particles are assumed as spheres and the contact handling is taken care.
This doesn't consider the shape of the surface of the rigid body in place.
Irregular bodies are modeled using many different variations now, one of the
methods is by overlapped sphere method \cite{das2007modeling}, SMR-DEM by
\citet{zhan2021surface}, dilated polyhedral DEM \citet{liu2020new}, Fourier
series-based discrete element method \citet{lai2020fourier}, non-smooth wall
modeling by \citet{amaro2019improvement}, GJK-DEM \cite{wachs2012grains3d},
discrete function representation based DEM \cite{lu2012critical}, level set
DEM method \cite{duriez2021precision}. However, these are complicated to
implement and not particle friendly and the rigid bodies have to discretized
differently to handle the non-smooth boundaries. In the current work we have
extended the contact force model proposed by \citet{mohseni2021particle} to
handle three dimensional problems, we also incorporate the damping model.

In the current work, we propose a new method to handle the collision between
arbitrary shaped rigid bodies. Unlike the approach of Mohseni, the proposed
contact force model is less sensitive to the primary secondary body selection.
And the proposed method is robust and very simple to implement when compared
with the existing techniques (cite zhang, Amaro) in the literature. The
current model tries to eliminate the confusion in choosing the primary and
secondary body in a collision instance. The rigid fluid coupling between CTVF
is new. We explored different rigid fluid coupling strategies by simulating
high density ratio simulations. Several examples are simulated to validate the
current scheme, ranging from simulations compared with FEM, analytical results
as well as experimental. Finally, in the interest of reproducibility and
easier ability for researchers to build on this work, our code is open source
and can be found at \url{https://gitlab.com/pypr/rfc}. We used automan package
\cite{automan2018} to automate all the results generated in the current
manuscript.

Numerical examples, collision of three cubes is initially used to test the
conservation property of the current contact mode while dealing with
non-smooth boundaries. Further, a free and forced sliding of a 3d cube on a
frictional inclined plane, and a rolling cylinder whose analytical solutions
exists are utilised to validate the rigid-rigid contact model frictional
model. A granular collapse of a stack of cylinders is modeled and compared
against the experimental results. In order to validate the two-way rigid fluid
coupled scheme a rigid circular body falling in a steady tank is simulated.
Complex interactions of fluid in a 3D dam breaking with multiple rigid bodies,
is studied and compared with the experimental results of SPH-DCDEM.

@@

@@

\subsection{Governing equations}
\label{sec:governing-eqns-fluid-dynamics}

SPH discretization of the modified governing equations are given by:
\begin{equation}
  \label{eq:sph-discretization-continuity}
  \frac{\tilde{d}\rho_a}{dt} = \sum_{b} \; \frac{m_b}{\rho_{b}} \; (
  \rho_{a} \; \tilde{\ten{u}}_{ab} \; + \;
  (\rho \; (\tilde{\ten{u}} \; - \;
  \ten{u}))_{ab}) \; \cdot \nabla_{a} W_{ab},
\end{equation}
\begin{multline}
  \label{eq:sph-momentum-fluid}
  \frac{\tilde{d}\ten{u}_{a}}{dt} = - \sum_{b} m_b \bigg[
  \bigg(\frac{p_a}{\rho_a^2} + \frac{p_b}{\rho_b^2}\bigg) \ten{I}
   \bigg]
  \cdot \nabla_{a} W_{ab} + \\
  \sum_{b} m_b \frac{4 \eta \nabla W_{ab}\cdot
    \ten{r}_{ab}}{(\rho_a + \rho_b) (r_{ab}^2 + 0.01 h_{ab}^2)} \ten{u}_{ab} +
  \ten{g}_{a},
\end{multline}

\newpage
~\newpage
~\newpage

@@

\newpage
~\newpage
The rigid body is discretized into particles with equal spacing each particle
with mass $m_i$ and density $\rho_i$. Rigid body has a total 6 degrees of
freedom (DOF), divided into $3$ translational and $3$ rotational.

The state of the rigid body at a given time (t) can described using position
and velocity of the center of mass ($\ten{x}_{cm}$), velocity of center of
mass ($\ten{v}_{cm}$), a rotation matrix ($\ten{R}$), to define the three
rotational degrees of freedom and angular velocity($\teng{\omega}$). The
center of mass is computed with
\begin{equation}
  \label{eq:center_of_mass}
  \ten{x}_{cm} = \frac{\sum_i m_i \; \ten{x}_{i} }{\sum_i m_i }
\end{equation}

The equations governing the dynamics of rigid body are, balance of linear and
angular momentum given by,
\begin{equation}
  \label{eq:balance_linear_mom}
  \frac{d \; (M \ten{v}_{cm})}{d t} = \sum_i \ten{F}_i,
\end{equation}
\begin{equation}
  \label{eq:balance_angular_mom}
  \frac{d \ten{L}}{d t} = \teng{\tau}_{cm}
\end{equation}
where $\teng{\tau}_{cm}$ is the torque about center of mass due to forces on the
particles,
\begin{equation}
  \label{eq:torque}
  \teng{\tau}_{cm} = \sum_i \ten{F}_i \times (\ten{r}_{cm} - \ten{r}_{i})
\end{equation}
and $\ten{L}$ is the angular momentum of the rigid body about center of mass,
computed as
\begin{equation}
  \label{eq:moi}
  \teng{L} =
  \sum_i \; \ten{r}_i \times \; (\ten{\omega} \times \ten{r}_i)
  = \sum_i \; m_i \; [(\ten{r} \cdot \ten{r}) \ten{I} - \ten{r} \otimes \ten{r}]
\end{equation}
\begin{equation}
  \label{eq:omega_compute}
  \teng{\omega} = \textit{\teng{I}}^{-1} \; \ten{L}
\end{equation}

The state of the rigid body at next time step is evolved through the
integration of governing \cref{eq:balance_linear_mom,eq:balance_angular_mom},
finding the linear velocity of the center of mass $\ten{v}_{cm}$ and angular
momentum $\ten{L}$ at the next timestep as,
\begin{equation}
  \label{eq:lin_pos_cm_update}
  \ten{x}_{cm}^{n+1} = \ten{x}_{cm}^{n} + \ten{v}_{cm}^{n} \; \Delta t
\end{equation}
\begin{equation}
  \label{eq:lin_vel_cm_update}
  \ten{v}_{cm}^{n+1} = \ten{v}_{cm}^{n} + \frac{\ten{F}_{cm}}{M} \; \Delta t
\end{equation}
 The orientation $\ten{R}$ is updated as,
\begin{equation}
  \label{eq:rotation_update}
  \ten{R}^{n+1} = \ten{R}^{n} + \tilde{\teng{\omega}}^{n} \, \ten{R}^{n} \; \Delta t
\end{equation}
\begin{equation}
  \label{eq:ang_mom_update}
  \ten{L}^{n+1} = \ten{L}^{n} + \teng{\tau}_{cm} \; \Delta t
\end{equation}
Moment of inertia at the new time step is computed as
\begin{equation}
  \label{eq:moi_update}
  (\textit{\teng{I}}^{-1})^{n+1} = \ten{R}^{n+1} \textit{\teng{\overline{I}}}^{-1} (\ten{R}^{n+1})^T
\end{equation}
The angular velocity at the new time step is computed with
\begin{equation}
  \label{eq:ang_velocity_update}
  \teng{\omega}^{n+1} = (\textit{\teng{I}}^{-1})^{n+1} \; \ten{L}^{n+1}.
\end{equation}
The position and velocity of the particles of the rigid body are updated by
\begin{equation}
  \label{eq:rb_particle_pos_update}
  \ten{r}_i = \ten{R} \cdot \overline{\ten{r}}_{i}
\end{equation}
\begin{equation}
  \label{eq:rb_particle_pos_update}
  \ten{x}_i = \ten{x}_{cm} + \ten{r}_{i}
\end{equation}
\begin{equation}
  \label{eq:rb_particle_vel_update}
  \ten{v}_i = \ten{v}_{cm} + \teng{\omega} \times \ten{r}_{i}
\end{equation}

% check \cite{natsui2018sph}
We use two coordinate frames to capture the dynamics of the rigid body. The
body fixed frame, which moves with rigid body is located always at the center
of mass ($\ten{x}_{cm}$) of the body, computed as

A body fixed coordinate system is put at the center of mass, and it translates
and rotates with the body. The discretized particle's position vector is
constant about this body fixed coordinate frame. Figure x shows the rigid body
at an instant, the position vector of particle $\ten{x}_i$ about global frame
is computed as
\begin{equation}
  \label{eq:global_position_vector_idea}
  \ten{x}_{i} = \ten{x}_{cm} + \ten{r}_i = \ten{x}_{cm} + \text{Rot}(\overline{\ten{r}}_i)
\end{equation}
where,
\begin{equation}
  \label{eq:body_position_vector}
  \overline{\ten{r}}_i = \ten{x}_i(0) - \ten{x}_{cm}(0)
\end{equation}

A 3$\times$3 rotation matrix is utilized to represent the orientation of the
body fixed coordinate frame, representing 3 degrees of freedom.
\begin{equation}
  \label{eq:rotation_matrix}
  \ten{A} = \mqty(\ten{a}_{1} & \ten{a}_{2} & \ten{a}_{3})
\end{equation}
where, each column $\ten{a}_{1}$, $\ten{a}_{1}$, $\ten{a}_{1}$ represents the
body frame coordinate axes unit vectors in global frame. With which the
position vector of the particle in local frame is translated to the global frame as,
\begin{equation}
  \label{eq:local_to_gloabl_translate}
  \ten{r}_i = \ten{A} \; \overline{\ten{r}}_i.
\end{equation}

\begin{figure}[!htpb]
  \centering
  % \includegraphics[width=1.0\textwidth]{images/contact_force/contact_force_description}
  \caption{2d description of local to global vector translation}
\label{fig:local_to_global_2d_explanation}
\end{figure}
Rotation matrix $\ten{A}$ has to be orthogonal, with determinant equal to $1$
and finally the magnitude of the eigenvalues is needed to be $1$. The position
of a body point $i$, can be written as
\begin{equation}
  \label{eq:global_position_vector}
  \ten{x}_{i} = \ten{x}_{cm} + \ten{A}(\overline{\ten{r}}_i)
\end{equation}

The state of the rigid body can be described by
\begin{equation}
  \label{eq:rotation_matrix}
  state = \mqty(\ten{x}_{cm} & \ten{A} & \ten{v}_{cm} & \ten{\omega})
\end{equation}
\newpage
~\newpage

@@

\begin{equation}
  \label{eq:force_on_fluid_due_to_rb}
  \frac{\tilde{d}\ten{u}_{a}}{dt} =  \sum_{b} m^0_b \bigg[
  \bigg(\frac{p_a}{\rho_a^2} + \frac{p_b}{\rho_b^2}\bigg) \ten{I}\bigg) \bigg]
  \cdot \nabla_{a} W_{ab}
\end{equation}

\begin{equation}
  \label{eq:force_on_fluid_due_to_rb}
  f_i =  m_a \sum_{b} m_b \bigg[
  \bigg(\frac{p_a}{\rho_a^2} + \frac{p_b}{\rho_b^2}\bigg) \ten{I}\bigg) \bigg]
  \cdot \nabla_{a} W_{ab}
\end{equation}

@@

\newpage
~\newpage
~\newpage

\FloatBarrier%
\section{Contact algorithm}
\label{sec:contact-algorithm}
\todoin{\citet{chen2019coupled} has damping model. Also he mentioned the parameters
required in stack of cylinders example.}

\begin{figure}[!htpb]
  \centering
  \includegraphics[width=1.0\textwidth]{images/contact_force/contact_force_description}
  \caption{Contact force description}
\label{fig:contact_foce_description}
\end{figure}
In the current work we have utilized the contact force model proposed by
\citet{mohseni2021particle}. The force acting on a particle $i$ of body A due
to the interaction with the particles of body $B$ can be resolved into a
normal and tangential component. Where, normal force component is utilised to
make sure that the particles of different body don't penetrate into each
other, while the tangential component is used to model the friction between
the interacting solids. According to \citet{mohseni2021particle}, we divide the
bodies under interaction into primary and secondary bodies, as shown in
\cref{fig:contact_foce_description}.
% In usual DEM to compute the force on particle i of body A, the force is
% computed by considering the overlap of particle i with each and every
% particle of body j by considering the particle j to be spherical. This leads
% to unphysical modeling of contact force when the body is interacting with a
% flat surface or surfaces which are not spherical by nature. The current
% contact force is surface aware. The force on particle $i$ is computed by
% equation
The normal force ($\teng{F}_i^{n}$) on
particle $i$ due to the interaction with the particles $j$ of body $B$ is
computed as,
\begin{equation}
  \label{eq:contact-algorithm-normal}
  \ten{F}_i = K_r \delta_{n}^{i} \ten{n}_i.
\end{equation}
Here, the overlap $\delta_{n}$ is computed using
\begin{equation}
  \label{eq:cf-distance-computation}
  d_i = \frac{
    \displaystyle\sum\limits_{j = 1}^{\text{NP}^{b}} \;
    \big( \ten{n}_i \cdot \ten{x}_{ij} \big)  \frac{m_j}{\rho_j} W_{ij}}
  {
    \displaystyle\sum\limits_{j = 1}^{\text{NP}^{b}} \;
    \frac{m_j}{\rho_j} W_{ij}}.
\end{equation}

The overlap is computed by $\delta_{n}^{i}$
\begin{equation}
  \label{eq:cf-overlap}
  \delta_{n}^{i} = \Delta x - d_i,
\end{equation}
where $\Delta x$ is the initial spacing between the particles. Note that while
computing the overlap of particle $i$ with the body $B$, we have computed an
effective overlap, rather than per particle interaction. This effectively is
able to model the interaction between non smooth surfaces, contrast from
particle particle force computation.

We associate a tangential spring attached to particle $i$ and body $B$ to
compute the tangential force, which initially has a magnitude of zero
($|\Delta \textit{\textbf{l}}_i|=0$). And the tangential spring is activated
when the particle comes into contact with body $B$, which is conformed by
\cref{eq:cf-overlap}. The tangential force is history-dependent. The contact
friction force is proportional to the tangential spring displacement, which is
integrated over the contact time as
\begin{equation}
  \label{eq:tangential-force}
  \ten{F}_{i}^{t^{n+1}} =
  -k_f \Delta \textit{\textbf{l}}_i^{\,n + 1} =
  -k_f \big[\big(\Delta {\textit{\textbf{l}}}_i^{\,n} \
  + \ten{v}_{ij}^{n + 1} \Delta t\big) \cdot \ten{t}_i^{n + 1} \big] \
  \ten{t}_i^{n + 1},
\end{equation}
where $\Delta t$ is the time step, $\ten{v}_{ij} = \ten{v}_{i} - \ten{v}_j$ is
the relative velocity of the primary particle $i$ with respect to the
secondary particle $j$. The tangential unit vector is computed by,
\begin{equation}
  \label{eq:tangential-vect}
  \ten{t}_i = \frac{\ten{v}_{ij} - (\ten{v}_{ij} \cdot \ten{n}_i) \ten{n}_i}{|\ten{v}_{ij} - (\ten{v}_{ij} \cdot \ten{n}_i) \ten{n}_i|}.
\end{equation}

The tangential force is coupled to the normal force through the Coulomb's law,
\begin{equation}
  \label{eq:Coulomb-law}
  \ten{F}_{i}^{t} = \min(\mu |\ten{F}_{i}^{n}|, |\ten{F}_{i}^{t}|) \
  \frac{\ten{F}_{i}^{t}}{|\ten{F}_{i}^{t}|}.
\end{equation}
This allows to impose the sliding friction condition between the interacting
solids. Finally, the total force acting on the particle $i$ due to the
interaction with body $B$ is:
\begin{equation}
  \label{eq:contact-force}
  \ten{F}_{i}^{cont} = \ten{F}_{i}^{n} + \ten{F}_{i}^{t}
\end{equation}

\begin{figure}[!htpb]
  \centering
  \includegraphics[width=0.3\textwidth]{images/contact_force/contact_force_description_3}
  \caption{Force transfer to the secondary particles $j$ from the primary body particle $i$}
\label{fig:secondary_particle_contact_foce_transfer}
\end{figure}
An equal and opposite force of the same magnitude is applied to the closest
secondary particle $j$ of $i$,
\begin{equation}
  \label{eq:contact-force}
  \ten{F}_{j}^{cont} = - \ten{F}_{i}^{cont}
\end{equation}

% The force on the particle $j$ is evaluated, on the
% body belonging to the secondary surface is as follows. Particles belonging to
% the secondary surface ($j$) which are at a distance less than the initial
% spacing ($l_0$) share the force exerted on particle i, as following:
% \begin{equation}
%   \label{eq:cf-overlap}
%   F^{ij} = - F^{i} w^{ij},
% \end{equation}
% where $w^{ij}$ is defined as
% \begin{equation}
%   \label{eq:cf-overlap}
%   w^{ij} = \frac{n^{i} \cdot \hat{\ten{r}}^{ij}}{\sum_j n^{i} \cdot \hat{\ten{r}}^{ij}}
% \end{equation}
